---
title: "Analytical Plan for Association between leadership commitment and professional development at NASA (2020): sex-adjusted stratified analysis"
subtitle: 'DOCUMENT: SAP-2022-008-GJ-v02'
author: '**From:** Felipe Figueiredo **To:** techsavvy32 (fiverr.com)'
date: "2022-02-02"
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAP_en_2.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', 160)
```

---

\newpage

**Analytical Plan for Association between leadership commitment and professional development at NASA (2020): sex-adjusted stratified analysis**

**Document version**

```{r history}
Version <- c("01", "02")
Changes <- c("Initial version", "Measure of effect added to the analysis, methods for inferential analyses were updated.")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

---

# Abbreviations

- CI: confidence interval
- FEVS: Federal Employee Viewpoint Survey
- OPM: U.S. Office of Personnel Management
- OR: odds ratio

# Context

The Federal Employee Viewpoint Survey (FEVS) addressed leadership commitment, professional development, and telework satisfaction while accounting for gender (OPM, 2020).
This analysis addresses a subset of the FEVS survey reflecting NASA employees.

## Objectives

Quantify the association between leadership commitment and
employee professional development
at NASA from the 2020 Federal Employee Viewpoint Survey.

## Hypotheses

There is a statistically significant correlation between leadership commitment, and professional development among NASAs employees in a telework environment after accounting for gender (men and women).

## Study design

Case-control to assess prevalence from a complex-designed survey, where sampling weights where used to account for sampling uncertainty.

# Data

```{r cleaning, include=FALSE}
source("scripts/input.R", encoding = 'UTF-8')
dv <- "(q1 -- I am given a real opportunity to improve my skills in my organization)"
iv <- "(q21 -- Supervisors in my work unit support employee development)"
```

## Raw data

```{r sample size}
# source population size
N <- data.raw %>% summarise(N_hat = sum(postwt)) %>% round
# source population size (cleaned)
N_hat <- analytical %>% summarise(N_hat = sum(postwt)) %>% round
```

Raw data was collected as a census of the eligible population (OPM, 2020), and statistical weighting was applied at the data collection to adjust for non-responses in in the census attempt.
These survey weights (variable `postwt`) allow for the estimation of the association under study in the source population.
The raw data was filtered to reflect only NASA survey respondents (where agency code equals `NN`).

The raw data is expected to reflect a total employee population at NASA at `r N` employees but after cleaning procedures the observations in the analytical data represents a total of `r N_hat` NASA employees.
Survey questions measured responses in a 5-point Likert scale between 1 (strongly disagree) and 5 (strongly agree).
Some questions offered the option to choose "X" (Don't know) as the answer.
These unknown answers were considered non-answers and treated as missing values (see section 5.1.4).

## Analytical dataset

All variables in the analytical set were labeled according to the raw data provided and values were labeled according to the data dictionary for the preparation of production-quality results tables and figures.

This analysis will focus on two questions from the FEVS survey, where the main interest is employee satisfaction
`r dv`
as the dependent variable and leadership commitment
`r iv`
as the independent variable.
As per the data cleaning process, the dependent variable was renamed to `dv` and the independent variable was renamed to `iv` in the analytical dataset.
Additionally, to calculate the OR the responses were categorized as binary responses, where agreement was aggregated from the "agree" and "strongly agree" responses, in variables `dv2` and `iv2`.

After the cleaning process `r ncol(analytical)` variables were included in the analysis with `r nrow(analytical)` observations.
Table 1 shows the structure of the analytical dataset.

```{r analytical data}
analytical_mockup %>%
  pander(caption = "**Table 1** Analytical dataset structure after variable selection and cleaning.")
```

The analytical dataset will be included in the private version of the report, and will be omitted from the public version of the report.

# Study variables

## Primary and secondary outcomes

**Specification of outcome measures** (Zarin, 2011):

1. (Domain) Employee professional development
2. (Specific measurement) Survey question
3. (Specific metric) End value
4. (Method of aggregation) Odds of participants that agree or strongly agree with survey question

**Primary outcome**

Odds of participants'
that perceive opportunities of employee professional development
at NASA from the 2020 Federal Employee Viewpoint Survey.

## Covariates

The association will be stratified by the sex of survey respondents.

# Statistical methods

## Statistical analyses

### Descriptive analyses

The epidemiological profile of the study participants will be described.
Demographic
(sex, age and BMI)
will be described as
mean (SD)
<!-- median (IQR) -->
or as counts and proportions (%), as appropriate.
The distributions of participants' characteristics will be summarized in tables and visualized in exploratory plots.

### Inferential analyses

All comparisons between groups will be performed as univariate analyses.
Differences in distribution of categorical variables will be assessed with the
chi-square test with the adjustment of design effect for weighted survey data.
The OR will be used as a measure of effect of the independent variable on the dependent variable.

The stratification by sex will be used to assess if the effect changes across male and female strata.
As a rule of thumb, a minimum change of 20% in the OR will be considered before concluding that there is an interaction between sex and the independent variable.
The homogeneity of the OR across strata will be assessed with the Cochran-Mantel-Haenszel test.

### Statistical modeling

N/A

### Missing data

No missing data imputation will be performed.
All evaluations will be performed as complete case analyses.

## Significance and Confidence Intervals

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be
two-tailed.

## Study size and Power

<!-- The G*Power software was used to determine the minimum sample size needed to achieve an 80% statistical power. -->
<!-- Using a Chi-squared statistical test with an alpha value of 5% for the sample size estimate, and an effect size of .5 (medium), the minimum sample size needed was 721. -->
<!-- From the completed surveys of 10,588, a random sampling method was selected to collect 721 completed and usable surveys from the 2020 OPM FEVS. -->
<!-- The researcher used all surveys to meet the inclusion criteria mentioned above to further increase the statistical power (Cohen et al., 2003). -->

N/A

## Statistical packages

This analysis will be performed using statistical software `R` version `r getRversion()`.

# Observations and limitations

N/A

# References

- **SAR-2022-008-GJ-v01** -- Association between leadership commitment and professional development at NASA (2020): sex-adjusted stratified analysis
- OPM (2020). 2020 Federal Employee Viewpoint Survey -- Technical report (<https://www.opm.gov/fevs/reports/technical-reports/>).
<!-- - Cohen, J. (1988). Statistical power analysis for the behavioral sciences (2nd Ed.). New York: Routledge. -->
- Zarin DA, et al. The ClinicalTrials.gov results database — update and key issues. N Engl J Med 2011;364:852-60 (<https://doi.org/10.1056/NEJMsa1012065>).
- Gamble C, et al. Guidelines for the Content of Statistical Analysis Plans in Clinical Trials. JAMA. 2017;318(23):2337–2343 (<https://doi.org/10.1001/jama.2017.18556>).

# Appendix

This document was elaborated following recommendations on the structure for Statistical Analysis Plans (Gamble, 2017) for better transparency and clarity.

## Availability

Both this analytical plan and the corresponding analysis report (**SAR-2022-008-GJ-v01**) can be downloaded in the following address:

<https://philsf-biostat.github.io/SAR-2022-008-GJ/>

<!-- The client has requested that this analysis be kept confidential. -->
<!-- Both this analytical plan and the corresponding analysis report (**SAR-2022-008-GJ-v01**) are therefore not published online and only the title and year of the analysis will be included in the consultant's Portfolio. -->
<!-- The portfolio is available at: -->

<!-- <https://philsf-biostat.github.io/> -->
